---
title: "Learning Machine Learning"
author: "Austin Halvorsen"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
  html_document:
    theme: flatly
    highlight: tango
    code_folding: hide
---

```{r load_library, include=FALSE}
# Use this R-Chunk to load your libraries!
library(mosaic)

```

```{r load_data, include=FALSE}
# Use this R-Chunk to load your datasets!

```

## Research Project: Deliverable {.tabset .tabset-fade .tabset-pills}

This project is a self research project for Business Intelligence and Analytics at Brigham Young Univerity - Idaho. The focus of this project has three parts:

1. Introduce Machine Learning and what it is
2. Give an example of machine learning on a sample dataset
3. Summarise my findings and share major learning points

Follow through the project using the tabs below. Also, feel free to connect with me on [LinkedIn](https://www.linkedin.com/in/ahalvo/) and follow me on [GitHub](http://www.github.com/austinhal)

### Project Introduction



### ML Example


### Conclusion


